The Galaxy Cluster Merger Catalog: An Online Repository of Mock Observations from Simulated Galaxy Cluster Mergers

### 2016-09-13 paper/ms.tex
Read through the paper and made some corrections/additions

\begin{abstract}
We present the first release of the ``Galaxy Cluster Merger Catalog''. This catalog provides an extensive suite of mock observations and related data for simulations of galaxy cluster mergers. These mock observations consist of projections of a number of important observable quantities in several different wavebands, for the entire evolution of each simulation as well as along different lines of sight through the three-dimensional simulation domain. The web interface to the catalog consists of easily browseable images over epoch and projection direction, as well as download links for the raw data and a JS9 interface for interactive data exploration. All of the data products are provided in the standard FITS file format, in image and table form. Future versions of the catalog will include simulations from a number of research groups and a variety of research topics related to the study of interactions of galaxy clusters with each other and with their member galaxies. The catalog is located at \url{http://gcmc.hub.yt}.
\end{abstract}

@@

The astrophysical literature is replete with simulations of galaxy cluster mergers, from binary merger simulations \citep[e.g.][]{ric01,poo06,zuh11,don13} to studies of mergers in cosmological simulations \citep[e.g.][]{vaz09,ski13,yu15}.

These simulations have often attempted to make predictions for what may be observed in a number of wavebands and made direct comparisons to observed merging systems. However, comparing the results of simulations of cluster mergers to these systems is often not straighforward; at what stage are we viewing the merger, and along what line of sight? To make matters more complicated, different combinations of merger epoch and line of sight can produce qualitatively similar projections of cluster emission, making it more difficult to determine these parameters.

We seek to address these issues by releasing the ``Galaxy Cluster Merger Catalog'', an extensive suite of mock observations and related data for simulations of galaxy cluster mergers. We have produced 2D projections and slices of a number of different quantities relevant to observations of galaxy clusters for a large number of merger simulations. The data products span the important stages of evolution of each merger and a number of relevant projection axes. The goal of this catalog is to provide a way to connect simulations of galaxy cluster mergers with multiwavelength observations of real merging clusters, providing an opportunity for observers to compare clusters in their observations with a particular merging scenario. The data may also be analyzed in its own right, to make predictions for what various investigations of real clusters may reveal for future observations and missions.

@@

\begin{itemize}
\item Each simulation is simulated using an adaptive or static mesh refinement (AMR/SMR) grid \citep{ber89} with varying resolution throughout the domain, with refinement occuring on criteria such as sharp jumps in density and temperature, matter density, and selected regions such as the cluster center. Eventually, the catalog will include particle-based simulations as well.
\item The equations of hydrodynamics (HD) or magnetohydrodynamics (MHD) are modeled using a conservative finite-volume scheme, employing Riemann solvers for evolving the flux of physical quantities and using high-order reconstruction schemes such as the Piecewise-Parabolic Method \citep[PPM,][]{col84}. Magnetic fields are evolved such that the condition $\nabla \cdot {\bf B} = 0$ is met, typically by a constrained transport scheme \citep[CT,][]{eva88}. Eventually, the catalog will include simulations peformed with smoothed particle hydrodynamics.
\item Each simulation assumes an ideal gas law equation of state with $\gamma = 5/3$ and primordial abundances of H/He with trace amounts of metals, yielding a mean molecular weight of $\mu = 0.6$.
\item If dark matter is included, it is modeled by an N-body solver for a collection of collisionless massive particles, which interact with the gas component only via gravity.
\item The gravity in the simulations is either modeled as a rigid gravitational potential associated with each cluster or by computing the self-gravity of the gas and dark matter using a Poisson solver \citep[e.g.,][]{ric08}.
\item Depending on the goals of the simulation study, other physics, such as viscosity, thermal conduction, radiative cooling, etc., may be included.
\end{itemize}

@@

\subsubsection{The yt Hub}\label{sec:yt_hub}

@@

The data products currently provided in the catalog are 2D selections or reductions of the original 3D datasets. The particular products included for a particular simulation depend on its physical and algorithmic details (e.g., slices of the magnetic field strength and Faraday rotation measure projections are only included for MHD simulations).

@@

With the exception of the X-ray event files (Section \ref{sec:xray}), no attempt is made in the mock observations to include statistical and systematic errors on the observed quantities, instrumental effects, or background contamination. It is not possible to anticipate what assumptions the end-user may want to make about these issues, so these effects have been left out so they may be added on a case-by-case basis for different situations.

@@

Projections are taken along several lines of sight. Currently, these include the three major axes of the simulation domain: $x$, $y$, and $z$. In the future, projections along off-axis directions will be added. For distance and redshift-dependent quantities, each simulation employs either the cosmology and redshift from the simulation, or for simulations without such information a $\Lambda{\rm CDM}$ cosmology is assumed with the following parameters: $H_0 = 71~{\rm km~s^{-1}~Mpc^{-1}},$ $\Omega_m = 0.27$, and $\Omega_\Lambda = 0.73$. The projection files each have a subset of the following fields contained within different HDUs with the following extension names:

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{epoch_page.png}
\caption{An example ``epoch page''. Links are provided to similar simulations with different physics at the same epoch. For each data type, figures of some of the fields are shown and links to download the data or display in JS9 are provided. This particular page is linked at \url{http://gcmc.hub.yt/virgo/novisc/0046/index.html}.}
\end{center}
\end{figure*}

Other fields related to the S-Z effect are also computed for some simulations, using the \code{SZpack}\footnote{\url{http://www.cita.utoronto.ca/~jchluba/Science_Jens/SZpack/SZpack.html}} library \citep{chl12, chl13} to compute the S-Z signal, including thermal and kinetic contributions as well as relativistic corrections. More details on how these projections were computed can be found at \url{http://yt-project.org/doc/analyzing/analysis_modules/sunyaev_zeldovich.html}. They are stored in separate FITS files from the other projections. Currently, these files contain a subset of these fields with the follwing extension names:

@@

\subsubsection{X-ray Events}\label{sec:xray}

The X-ray emissivity projection listed above represents a single broad energy band from 0.5-7.0 keV. To do analysis of the thermal and compositional properties of the ICM, high-resolution energy spectra are required. Instead of providing maps at the same spatial resolution for hundreds or even thousands of energy bins for every epoch and projection direction, which would be prohibitive in terms of disk space, the catalog provides event files.

The X-ray event files are standard products which can be manipulated and analyzed with common X-ray analysis tools, such as ds9, \code{CIAO}, and the \code{HEASOFT} software suite. The events themselves have been generated from the 3-D datasets using the \code{pyXSIM}\footnote{\url{http://hea-www.cfa.harvard.edu/~jzuhone/pyxsim}} package \citep{zuh14}. At the time of writing, event files for the {\it Chandra} ACIS-I detector are provided, but eventually event files for other instruments will be included, as well as raw, unconvolved events which may be used with any instrument simulator.

The event files are designed to provide approximate representations of real observations, with some simplifications. First, the event energies have been convolved with the ACIS-I on-axis responses over the entire field of view. Second, the pixel size corresponds to the width of the finest simulation cell size (in line with the other data products), instead of the pixel scale of the detector. Thirdly, since for the distances assumed here the cell sizes are much larger than the {\it Chandra} PSF, no PSF effects have been applied. The idea is not to create an exact simulation, but instead to approximate the statistical properties of the source while allowing one to use standard X-ray tools to analyze the spectral and spatial properties of the data.

\subsubsection{Galaxies}\label{sec:galaxies}

A number of the simulations have been run using an N-body code to represent the collisionless dark matter component of the cluster using massive particles. A number of these dark matter particles can be sampled from the simulation to represent mock ``galaxies'', which can also be treated as a collisionless component. For some of these simulations, we have created mock galaxy catalogs.

The number of particles that have been drawn per halo is determined by the mass-richness relation of \citet{for14}. The FITS binary tables for the galaxies include sky positions, line-of-sight velocities, an identifier for which halo each galaxy originally belonged to, and unique IDs for every galaxy. These particles provide a way of measuring the kinematics of the merger from the perspective of the collisionless material with a statistical significance that is comparable to that obtained from measured redshifts of galaxies in real clusters. No redshift errors have been applied to the galaxy velocities, which is an exercise left to the end-user.

A ds9 region file is provided for each epoch and line-of-sight in addition to the FITS file, which allows the galaxy positions to be plotted over the projections of the other fields.

@@

\acknowledgements
JAZ thanks Eric Mandel for tips on using JS9 and for critical bug fixes and enhancements that were provided to make it work with the data in the catalog. JAZ acknowledges support through Chandra Award Number G04-15088X issued by the Chandra X-ray Center, which is operated by the Smithsonian Astrophysical Observatory for and on behalf of NASA under contract NAS8-03060. This research made use of AstroPy, a community-developed core Python package for Astronomy \citep{apy13}.

>>>>

\begin{abstract}
We present the first release of the ``Galaxy Cluster Merger Catalog''. This catalog provides an extensive suite of mock observations and related data for N-body and hydrodynamical simulations of galaxy cluster mergers. These mock observations consist of projections of a number of important observable quantities in several different wavebands, for the entire evolution of each simulation as well as along different lines of sight through the three-dimensional simulation domain. The web interface to the catalog consists of easily browseable images over epoch and projection direction, as well as download links for the raw data and a JS9 interface for interactive data exploration. All of the data products are provided in the standard FITS file format, in image and table form. Future versions of the catalog will include simulations from a number of research groups and a variety of research topics related to the study of interactions of galaxy clusters with each other and with their member galaxies. The catalog is located at \url{http://gcmc.hub.yt}.
\end{abstract}

@@

The astrophysical literature is replete with simulations of galaxy cluster mergers, from binary merger simulations \citep[e.g.][]{ric01,poo06,zuh11,don13} to studies of mergers in cosmological simulations \citep[e.g.][]{vaz09,ski13,yu15}. These simulations have often attempted to make predictions for what may be observed in a number of wavebands and made direct comparisons to observed merging systems. However, comparing the results of simulations of cluster mergers to these systems is often not straighforward; at what stage are we viewing the merger, and along what line of sight? To make matters more complicated, different combinations of merger epoch and line of sight can produce qualitatively similar projections of cluster emission, making it more difficult to use observations to determine these parameters and distinguish between different theoretical models.

We seek to address these issues by releasing the ``Galaxy Cluster Merger Catalog'', an extensive suite of mock observations and related data for simulations of galaxy cluster mergers. We have produced 2D projections and slices of a number of different quantities relevant to observations of galaxy clusters for a large number of merger simulations. The data products span the important stages of evolution of each merger and a number of relevant projection axes. The goal of this catalog is to provide a way to connect simulations of galaxy cluster mergers with multiwavelength observations of real merging clusters, providing an opportunity for observers to compare clusters in their observations with a particular merging scenario. The data may also be analyzed to make predictions for what various investigations of real clusters may reveal for future observations and missions.

@@

\begin{itemize}
\item Each simulation is simulated using an adaptive or static mesh refinement (AMR/SMR) grid \citep{ber89} with varying resolution throughout the domain, with refinement occuring on criteria such as sharp jumps in density and temperature, matter density, and selected regions such as the cluster center.
\item The equations of hydrodynamics (HD) or magnetohydrodynamics (MHD) are modeled using a conservative finite-volume scheme, employing Riemann solvers for evolving the flux of physical quantities and using high-order reconstruction schemes such as the Piecewise-Parabolic Method \citep[PPM,][]{col84}. If present, magnetic fields are evolved such that the condition $\nabla \cdot {\bf B} = 0$ is met, typically by a constrained transport scheme \citep[CT,][]{eva88}.
\item Each simulation assumes an ideal gas law equation of state with $\gamma = 5/3$ and primordial abundances of H/He with trace amounts of metals, yielding a mean molecular weight of $\mu \simeq 0.6$.
\item If dark matter is included, it is modeled by an N-body solver for a collection of collisionless massive particles, which interact with the gas component only via gravity.
\item The gravity in the simulations is either modeled as a rigid gravitational potential associated with each cluster or by computing the self-gravity of the gas and dark matter using a Poisson solver \citep[e.g.,][]{ric08}.
\item Depending on the goals of the simulation study, other physics, such as viscosity, thermal conduction, radiative cooling, etc., may be included.
\end{itemize}

Though currently the simulations represented in the catalog are of the grid-based variety, eventually the catalog will also include data from smoothed particle hydrodynamics (SPH) simulations.

@@

\subsubsection{The yt Hub}\label{sec:yt_hub}
The catalog data is stored on the yt Hub~(\textbf{citeme}), a conglomerate of various open source services creating an environment that allows one to collect, remotely analyze, share and publish scientific datasets. At its core, the yt Hub uses Girder~(\textbf{citeme}), a data management platform that allows to transparently store, serve, and proxy data from heterogeneous backend storage engines through a single web API. Data served by Girder can easily be gathered into dynamical hierarchies such as collections, folders and items. This data management system is integrated with JupyterHub~(\textbf{citeme}) and allows authenticated users to spawn Jupyter notebook servers with direct access to the data that is provided as a dynamically composed FUSE filesystem. Any Jupyter notebook created in that environment during an interactive session is safely stored back in the Girder instance, preserving the user's analysis workflow and increasing research reproducibility. In the future, we plan to have this notebook capability integrated directly with the catalog website.

@@

The data products currently provided in the catalog are 2D selections or reductions of the original 3D datasets, such as slices or projections. The particular products included for a particular simulation depend on its physical and algorithmic details (e.g., slices of the magnetic field strength and Faraday rotation measure projections are only included for MHD simulations).

@@

With the exception of the X-ray event files (Section \ref{sec:xray}), no attempt is made in the mock observations to include statistical and systematic errors on the observed quantities, instrumental effects, or background contamination. It is not possible to anticipate in advance what assumptions the end-user may want to make about these issues, so these effects have been left out so they may be added on a case-by-case basis for different situations.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{epoch_page.png}
\caption{An example ``epoch page''. Links are provided to similar simulations with different physics at the same epoch at the top of the page. For each data type, figures of some of the fields are shown and links to download the data or display in JS9 are provided. This particular page is linked at \url{http://gcmc.hub.yt/virgo/novisc/0046/index.html}.}
\end{center}
\end{figure*}

Projections are taken along several lines of sight. Currently, these include the three major axes of the simulation domain: $x$, $y$, and $z$. In the future, projections along off-axis directions will be added. For distance and redshift-dependent quantities, each simulation employs either the cosmology and redshift from the simulation, or for simulations without such information, a redshift and a $\Lambda{\rm CDM}$ cosmology is assumed with the following parameters: $H_0 = 71~{\rm km~s^{-1}~Mpc^{-1}},$ $\Omega_m = 0.27$, and $\Omega_\Lambda = 0.73$. The projection files each have a subset of the following fields contained within different HDUs with the following extension names:

@@

Projections of the full S-Z signal are also computed for some simulations, using the \code{SZpack}\footnote{\url{http://www.cita.utoronto.ca/~jchluba/Science_Jens/SZpack/SZpack.html}} library \citep{chl12, chl13} to compute the S-Z signal, including thermal and kinetic contributions as well as relativistic corrections. More details on how these projections were computed can be found at \url{http://yt-project.org/doc/analyzing/analysis_modules/sunyaev_zeldovich.html}. They are stored in separate FITS files from the other projections. Currently, these files contain a subset of these fields with the follwing extension names:

@@

\subsubsection{X-ray Event Files}\label{sec:xray}

The X-ray emissivity projection listed above represents a single broad energy band from 0.5-7.0 keV. However, to do analysis of the thermal and compositional properties of the ICM, high-resolution energy spectra are required. Instead of providing similar maps at the same spatial resolution for hundreds or even thousands of energy bins for every epoch and projection direction, which would be prohibitive in terms of disk space, the catalog provides event files.

These X-ray event files are standard products which can be manipulated and analyzed with common X-ray analysis tools, such as ds9, \code{CIAO}, and the \code{HEASOFT} software suite. The events themselves have been generated from the original 3-D datasets using the \code{pyXSIM}\footnote{\url{http://hea-www.cfa.harvard.edu/~jzuhone/pyxsim}} package \citep{zuh14}, which not only generates the 2-D projected X-ray events but can also convolve the photons with instrumental response matrices. At the time of writing, event files for the {\it Chandra} ACIS-I detector are provided, but eventually event files for other instruments will be included, as well as raw, unconvolved events which may be used as input for any instrument simulator.

The event files are designed to provide approximate representations of real observations, with some simplifications. First, the event energies have been convolved with the ACIS-I on-axis responses over the entire field of view. Second, the pixel size corresponds to the width of the finest simulation cell size (in line with the other data products), instead of the pixel scale of the detector. Thirdly, since the {\it Chandra} PSF is very small and for the distances assumed here the cell sizes are typically much larger, no PSF effects have been applied. The idea is not to create an exact simulation, but instead to approximate the statistical properties of the source while allowing one to use standard X-ray tools to analyze the spectral and spatial properties of the data.

\subsubsection{Mock Galaxy Catalogs}\label{sec:galaxies}

A number of the simulations have been run using an N-body code to represent the collisionless dark matter component of the cluster via massive particles. A number of these dark matter particles can be sampled from the simulation to represent mock ``galaxies'', which can also be treated as a collisionless component. For some of these simulations, we have created mock galaxy catalogs.

A number of particles are drawn from each halo within a radius of $r_{200}$, determined by the mass-richness relation of \citet{for14}, from the intitial simulation epoch, which are the ``galaxies''. The same galaxy particles are tracked for the rest of the epochs in the simulation. The FITS binary tables for the galaxies include sky positions, line-of-sight velocities, an identifier for which halo each galaxy originally belonged to, and unique IDs for every galaxy. These particles provide a way of measuring the kinematics of the merger from the perspective of the collisionless material with a statistical significance that is comparable to that obtained from measured redshifts of member galaxies in real clusters. No redshift errors have been applied to the galaxy velocities, which is an exercise left to the end-user.

A ds9 region file containing point regions corresponding to the galaxy positions is provided for each epoch and line-of-sight in addition to the FITS file, which allows the galaxy positions to be plotted over the projections of the other fields.

@@

\acknowledgements
JAZ thanks Eric Mandel for tips on using JS9 and for critical bug fixes and enhancements that were provided to make it work with the data in the catalog. JAZ acknowledges support through Chandra Award Number G04-15088X issued by the Chandra X-ray Center, which is operated by the Smithsonian Astrophysical Observatory for and on behalf of NASA under contract NAS8-03060.
